---
title: "Covid91 vaccine study Final2021"
author: "Angel Ramos Mendiola"
date: "May 4th, 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods 
In order to find the drug's effectiveness on the different genders (Males and Females) and circumstances that the people are in (LGBTQ and Drug Users), I am going to first isolate Male, Female, LGBTQ, and Drug Users to ultimately get a accurate result for all 4. 

```{r}
maledata <-filter(FinalData,sex=="guy") 
maledata$sex <- as.factor(maledata$sex)  
maledata$infected =factor(maledata$infected)
maledata$infected <- relevel(maledata$infected, "OK")
```

```{r}
maledata <-filter(FinalData,sex=="guy") 
```

```{r}
femaledata <-filter(FinalData,sex=="gal") 
femaledata$sex <- as.factor(femaledata$sex)
femaledata$infected =factor(femaledata$infected)
femaledata$infected <- relevel(femaledata$infected, "OK")
levels(femaledata$infected)
```

```{r}
femaledata <-filter(FinalData,sex=="gal") 
femaledata$sex <- as.factor(femaledata$sex)
```

```{r}
druggiesdata <- filter(FinalData, DrugUser == "yes")
druggiesdata$DrugUser <- as.factor(druggiesdata$DrugUser)
druggiesdata$infected =factor(druggiesdata$infected)
druggiesdata$infected <- relevel(druggiesdata$infected, "OK")
```

```{r}
LGBTQdata <-filter(FinalData,LGBTQ=="gay") 
LGBTQdata$sex <- as.factor(LGBTQdata$LGBTQ)
LGBTQdata$infected =factor(LGBTQdata$infected)
LGBTQdata$infected <- relevel(LGBTQdata$infected, "OK")
```

# Males 

## Bar Chart
```{r}
Males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

After reviewing the Bar Chart, I can estimate that the percentage of the 340 males that participated in the study who were infected with Covid-19 that got the drug was around 58.5%, now the percentage of the males who were infected with Covid-19 that got the placebo was around 41.5%. 

## Numerical Summary
```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

After reviewing this Numerical Chart, around 340 males that were in the study caught Covid-19 out of the 10,787 men tested. The table shows me that 3.72% of males who took the drug caught Covid-19 and the 2.55% of the men that got injected with the placebo caught Covid-19. The Numerical Summary shows that men are 50% more likely to catch Covid-19 after being injected with the vaccine, this shows that the vaccine is not effective. 

## Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

After seeing the Inferential Results, I can see that the P-Value for both the Fisher Exact Test (0.0006181) and the Chi-squared (0.0005801) are relatively small. This can tell me that the results you see on the Inferential Results are accurate and not by any coincidence.   
$H_0$ The Null Hypothesis shows that there is not a difference in men catching Covid-19 if they get vaccine or not.  
$H_A$ The Alternative Hypothesis shows that the vaccine does in fact make a difference in men catching Covid-19 if they do get the vaccine or not. So, I reject the Null Hypothesis after seeing the Alternative Hypothesis. 

The odds ratio from the Fisher Exact test is 1.47:1, when taking the odds ratio and computing it into a percentage of the effectiveness  you can do this by using the formula 0/(1+0)*100 you get a 0.5951417% effectiveness.   

The vaccine is 40% effective for the men, which means that the vaccine is ineffective.
```
```


# Females

## Bar Chart
```{r}
Females <- femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```



## Numerical Summary 
```{r}
table2 <- xtabs(~infected + treatment, data=femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```

## Inferential Results
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The vaccine is 67% effective in females. 


# Drug Users

## Bar Chart
```{r}
Druggies <- druggiesdata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```




## Numerical Summary 
```{r}
table3 <- xtabs(~infected + treatment, data=druggiesdata)
rowPerc(table3)
colPerc(table3)
print(table3)
```

## Inferential Results
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```
The vaccine is 60% effective in the Drug User population. 

# LGBTQ

## Bar Chart
```{r}
LGBTQ <- LGBTQdata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

## Numerical Summary 
```{r}
table4 <- xtabs(~infected + treatment, data=LGBTQdata)
rowPerc(table4)
colPerc(table4)
print(table4)
```

## Inferential Results
```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The vaccine is not effective in the LGBTQ community, seeing the fact that it is only 29% effective.

# Discussion and Conclusion 

After looking at the 4 different groups in the study

